Stateless Model Checking with Data-Race Preemption Points

### 2015-10-16 pldi-16/di.tex
Do some more stuff and write about portend

@@

{\bf Terminology.}
Importantly, we distinguish data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is false-positive-prone, Landslide reports unprotected access pairs separately from the problems listed above, calling the pair a {\em data race candidate}.
Should a future interleaving, preempting threads during those accesses, expose e.g. a crash or deadlock, then we report a {\em data-race bug}.

@@

>>>>

{\bf Terminology.}
Importantly, we distinguish data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is false-positive-prone, Landslide reports unprotected access pairs separately from the problems listed above, calling the pair a {\em data race candidate}.
Should a future interleaving, preempting during those accesses, cause a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
For brevity, we classify benign data races as a subset of false positives.

@@

@@

% If there's room, mention the cant_swap mechanism for killing the top half of deferred jobs.

### 2015-10-16 pldi-16/evaluation.tex
Do some more stuff and write about portend

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

@@

{\bf Avoiding false negative data-race bugs.}
Some pairs of unprotected shared-memory accesses may be hidden in some control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass data-race detector
%could not achieve the coverage necessary
would fail
to identify the access pair as a data-race candidate to begin with.
%
We counted how many such data-races, used as PPs, would lead to \landslide~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
% TODO: Put a figure here giving an example of where e.g. a data race only shows up during the contention path of a mutex.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).
%
To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported free-re-malloc data races during the first interleaving as {\em deterministic}, as a single-pass analysis must, rather than waiting until future interleavings to confirm them (see \sect{\ref{sec:free-re-malloc}}).

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% TODO: Argue:
				% It is fair to compare multiple pass DR-analysis under Landslide against just a single execution because prior work DR detectors, being not integrated with a MC, are not intended to uncover different results under subsequent runs.
				% Define a "stress tester" as a class of bug detectors where they are intended to [...]
				% Maybe say: Do there exist any stress-tester DR detectors where they are intended to produce new results under reruns?

				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

@@

{\bf Avoiding false negative data-race bugs.}
\label{sec:eval-falseneg}
Some pairs of unprotected shared-memory accesses may be hidden in some control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass data-race detector
%could not achieve the coverage necessary
would fail
to identify the access pair as a data-race candidate to begin with.
%
We counted how many such data-races, used as PPs, would lead to \landslide~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
% TODO: Put a figure here giving an example of where e.g. a data race only shows up during the contention path of a mutex.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).
%
To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported free-re-malloc data races during the first interleaving as {\em deterministic}, as a single-pass analysis must, rather than waiting until future interleavings to confirm them (see \sect{\ref{sec:free-re-malloc}}).

### 2015-10-16 pldi-16/related.tex
Do some more stuff and write about portend

%None. We are the first ever to explore this area, and of course expect the Turing Award for our groundbreaking insights.

\quicksand~is built upon many established concurrency testing techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%Our model checker
\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related model checkers largely based on their treatment of shared-memory thread communication.

@@

@@

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs. However, every such prior tool we know of has serious drawbacks.
{\textsc SKI} \cite{ski} approaches kernel code by choosing in advance a random set of instruction offsets from the start of the test,
which is more similar to stress testing or fuzzing than to exhaustive state space exploration.
% TODO: Make sure you talk about your mutex experiment, or rephrase this.
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU, which is very similar to our \mxtest~experiment.
However, it does not employ Iterative Deepening, and requires programs to be written in the PROMELA DSL.
%so cannot check implementations directly.

@@

blah blah blah eraser

@@

>>>>

% TODO: ", and related data race analyses based on ????"

@@

We build upon many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related model checkers by their treatment of shared-memory thread communication.

@@

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs. However, every such prior tool we know of has serious drawbacks.
% TODO: Be proud of them instead of being dismissive.
{\textsc SKI} \cite{ski} approaches kernel code by choosing in advance a random set of instruction offsets from the start of the test,
which is more similar to stress testing or fuzzing than to exhaustive state space exploration.
% TODO: Make sure you talk about your mutex experiment, or rephrase this.
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU, which is very similar to our \mxtest~experiment.
However, it does not employ Iterative Deepening, and requires programs to be written in the PROMELA DSL.
%so cannot check implementations directly.

@@

@@

\newcommand\portend{\textsc{Portend}}
\portend~\cite{portend} is the most closely related work we have found.
It performs a single-pass analysis as normal, then explores a limited state space to classify candidates in a taxonomy of likely severity.
Compared to us, \portend~can detect non-failing races which nevertheless cause suspiciously different program output,
while we depend on directly detecting failures.
It also uses symbolic execution to test input nondeterminism along with schedule nondeterminism,
while we explore the latter only.
However, \portend~does not perform MC-like exploration in advance of knowing data-race locations,
which we showed in \sect{\ref{sec:eval-falseneg}} can miss some bugs.
It is also limited to the POSIX synchronization API, so cannot verify lock implementations such as we do with \mxtest.
Future work could combine the two approaches, using \landslide~to produce new traces for \portend~to classify, or using \portend's analysis to inform \quicksand's state space priorities.
